WALi-NWA is a C++ library for constructing, querying, and operating on
nested-word automata.  It is a portion of the WALi library, which provides
types and operations for weighted automata.
While the NWA portions of WALi
are mostly logically separate from the rest of WALi, it does use
facilities provided by WALi and inter-operates with WALi's weighted pushdown
system (WPDS) code.

We assume the reader is familiar with nested words and
nested-word automata; if not, \appref{nwa-definition} provides the definition
we use and
Alur et al.~\cite{DLT:AM2006,JACM:AM2009} describe them in detail. Our
implementation corresponds to the definition from the DLT 2006
paper~\cite{DLT:AM2006}; relative to followup work~\cite{JACM:AM2009}, this definition removes
the distinction between the machine's linear and hierarchical states, and on
a call transition always stacks the source state. In addition, we only require that the
final linear state be accepting for an NWA to accept a
word. (Following~\cite{JACM:AM2009}, this specific variant is a
``weakly-hierarchical, linearly-accepting'' NWA.)

The terminology used in the interface of the package is geared
toward program analysis, so we use terms such as ``call site'' and ``return
site'' to refer to states of the NWA even though they take on different
meanings in different contexts.

In contrast to the standard definitions, we allow internal $\epsilon$
transitions (but not calls or returns). We also allow ``wild''
transitions, which match any single symbol. (Wilds \emph{can} appear on calls
and returns.)

%The NWAs in this library have support for balanced and unbalanced-left nested
%words only, and does not provide support for any words with pending
%returns. (That is, we support balanced words and nested-word prefixes, but
%not nested-word suffixes or general nested words.)

The WALi library is available from
\url{http://www.cs.wisc.edu/wpis/wpds/download.php}. This document describes
the NWA portion of version 4.0. Building instructions can be found in
\texttt{README.txt} in the root directory of the distribution.

\vfill

{\noindent\footnotesize This document generated from WALi repository revision \input{|svnversion}}

\clearpage
\tableofcontents
\clearpage

\section{Library Overview}
\label{Se:Nested Word Automata}

The core of the NWA library is in the namespace \texttt{wali::nwa}. It
includes the classes \texttt{NestedWord}, \texttt{NWA}, and others.

There are a large number of non-member functions that operate on NWAs. These
are divided into three sub-namespaces:
\texttt{wali::nwa::query}, \texttt{wali::nwa::construct}, and
\texttt{wali::nwa::nwa\_pds}. There is also an experimental parser in
\texttt{wali::nwa}.

Finally, there are some operations and classes provided by WALi that the NWA
portion of the library uses. This mostly consists of the key-handling code,
but can also include the WPDS portions of the library.

Throughout this document, include paths are relative to the \texttt{Source}
directory from the WALi distribution.

Section \ref{Se:ExampleUse} illustrates a simple sample use of the library.

\subsection{NWA core classes}

These types live in namespace \texttt{wali::nwa}:

\begin{functionlist}
  \functionitem[\texttt{NestedWord}] Implements a single nested word. Currently the only
    operation that can be performed using one is checking whether a
    \texttt{NestedWord} is a member of an \texttt{NWA}'s language. Defined in
    \texttt{wali/nwa/NestedWord.hpp}. See \sectref{class-nested-word}.
  \functionitem[\texttt{NWA}] Implements a nested-word automaton. Defined in
    \texttt{wali/nwa/NWA.hpp} with a forward declaration in \texttt{wali/nwa/NWAFwd.hpp}.
    See
    \sectref{NWA-class}. (\sectref{serialization} also discusses member
    functions related to serializing an NWA.)
  \functionitem[\texttt{NWARefPtr}] \texttt{NWARefPtr} is a typedef of \texttt{ref\_ptr$<$NWA$>$} (see
    below). This is defined in \texttt{wali/nwa/NWAFwd.hpp}.
  \functionitem[\texttt{State} and \texttt{Symbol}] These are both typedefs of
    \texttt{wali::Key} (see below). Both are defined in
    \texttt{wali/nwa/NWAFwd.hpp}. \emph{Note:} the use of keys for both
    states and symbols means that they can be confused from a types
    perspective, so use caution that this does not happen. 
  \functionitem[\texttt{StateSet} and \texttt{SymbolSet}] These are typedefs of
    \texttt{std::set}s of the corresponding type. (Client code should not
    rely on this
    fact; it could change to be an \texttt{unordered\_map} or other similar
    container.) Both are defined in \texttt{wali/nwa/NWAFwd.hpp}.
  \functionitem[\texttt{ClientInfo}] This class holds client-specific information that
    is associated with a state. Client code can subclass \texttt{ClientInfo} and use
    functions of the \texttt{NWA} class to attach instances to states. To use,
    include \texttt{wali/nwa/ClientInfo.hpp}. See \sectref{client-info}.
  \functionitem[\texttt{WeightGen}] This abstract class describes how to assign weights
    to transitions of the NWA, and is used both when converting an NWA to a
    PDS and when doing prestar/poststar queries on the NWA directly. To use,
    include \texttt{wali/nwa/WeightGen.hpp}. See \sectref{NWAtoPDS}.
\end{functionlist}


\subsection{NWA non-member functions}

The library provides a large number of non-member functions for operating on
NWAs. These are partitioned into the following namespaces:

\begin{functionlist}
  \functionitem[\texttt{wali::nwa::query}] This namespace provides functions for
    querying an automaton. The kinds of functions in this namespace are:
    \begin{itemize}
      \item Functions that query transitions of
        the NWA, returning information about one of the states or the symbol
        in transitions that meet some criteria. (For example, ``give me all
        states that appear as the target state of an internal transition with
        this source state.'') See \sectref{query-transitions}.
      \item Determining whether two NWAs have any states in
        common. See \sectref{query-automaton}.
      \item Determining whether an \texttt{NWA} is deterministic. See
        \sectref{query-automaton}.
      \item Determining whether a \texttt{NestedWord} is a member of the
        language of an NWA. See \sectref{query-language}.
      \item Determining whether the language of an NWA is empty. See
        \sectref{query-language}.
      \item Determining whether the languages of two NWAs are equal, or if
        one is a subset of the other. See \sectref{query-language}.
      \item prestar and poststar operations on an NWA. See
        \sectref{query-language}.
    \end{itemize}
    
  \functionitem[\texttt{wali::nwa::construct}] This namespace provides functions for
    constructing an NWA. Functions in this namespace are standard,
    automata-theoretic operations such as intersection, union, concatenation,
    Kleene star, and reversal. See \sectref{Building NWAs}.
  \functionitem[\texttt{wali::nwa::nwa\_pds}] This namespace provides functions for
    converting between NWAs and PDSs, and related functions. (Note that
    construction of an NWA from a PDS is in this namespace, not in
    \texttt{construct}.) See \sectref{Conversions}.
  \functionitem[\texttt{wali::nwa}] This namespace, in addition to the items already
    discussed, provides an experimental parser for a serialization format of
    NWAs. See \sectref{serialization}.
\end{functionlist}


\subsection{Generic WALi}

The NWA library uses these portions of the standard WALi library. Unless
otherwise specified, they are in namespace \texttt{wali}. See \cite{wali}.

\begin{functionlist}
  \functionitem[\texttt{ref\_ptr$<$T$>$}] This is a reference-counted, intrusive
    smart-pointer template. (``Intrusive'' means that the pointed-to type
    must be modified to include a \texttt{count} field used by the
    pointer. This is most conveniently done by subclassing
    \texttt{wali::Countable}.) Defined in \texttt{wali/ref\_ptr.hpp}.

  \functionitem[\texttt{Key}] A \texttt{Key} is a unique identifier naming states and symbols in an
    NWA or NestedWord. It is actually a typedef of an integer, though client
    code should not rely on this precise type. Declared in
    \texttt{wali/Key.hpp}.

  \functionDef{Key}{getKey}{...}{} This function produces a key from some input. If
    the input has not been seen before, it returns a new key; if it has, then
    it returns the same key as before. (One can think of this as translating
    whatever unique identifier is known by the client code to the
    \texttt{Key} needed by WALi.) Overloads of this function are provided for
    the following types:
    \begin{itemize}
      \item \texttt{std::string const \&}
      \item \texttt{char *}
      \item \texttt{int}
      \item \texttt{std::set<Key> const \&}
      \item \texttt{Key, Key} (this is a two-argument version of
        \texttt{getKey})
      \item \texttt{key\_src\_t}
    \end{itemize}
    The final overload, for \texttt{key\_src\_t}, is a \texttt{ref\_ptr} to a
    \texttt{KeySource} object. \texttt{KeySource} is an abstract class that
    client code can subclass to provide a translation to \texttt{Key}s for
    arbitrary types.

    All overloads are declared in \texttt{wali/Key.hpp}.

  \functionDef{std::string}{key2str}{Key k}{}
    \texttt{key2str} is an ``inverse'' of \texttt{getKey}, except that it
    always returns a string representation. If the key was created with
    \texttt{getKey(std::string const \&)} in the first place, this
    returns a copy of the original string.\footnote{For \texttt{int}s, it is the
    textual representation of the original number. For a \texttt{set<Key>},
    it is the set printed in standard set notation. For a pair of keys, it is
    the pair printed as an ordered pair. In general, it is the result of
    calling \texttt{key\_src\_t::print}. See the \texttt{print} function in
    each of the files \texttt{wali/*Source.cpp}.} Declared in
    \texttt{wali/Key.hpp}.

  \functionitem[\texttt{wali::wfa::WFA}] This is a weighted finite automaton. It is
    used in NWA poststar and prestar queries in much the same manner as in
    WPDS poststar and prestar queries. Defined in \texttt{wali/wfa/WFA.hpp}.

  \functionitem[\texttt{wali::wpds::WPDS}] This is a weighted pushdown system. NWAs use
    WPDSs behind the scenes when doing poststar and prestar queries, and the
    library provides conversion routines between the two automata
    types. Defined in \texttt{wali/wpds/WPDS.hpp}.
\end{functionlist}


\subsection{Simple Example use of the Library}
\label{Se:ExampleUse}
The following is an illustration of basic operations of the library. The full
code is available in the file \texttt{Doc/NWA\_tex/Example/example.cpp} in the
WALi distribution.

\inputminted{c++}{Example/example-abbrev.cpp}

